Hierarchical Large Language Models for Multi-task problems

### 2024-01-14 paper/main.tex
draft a bit in the paper

@@

In addition to these two approaches, we present generalization of the two approaches: the Hierarchical Low-Rank Adaption.
Given a pretrained LLM and $n$ tasks, we construct $n$ sets of LoRA parameters $\theta_n$ that each are trained on its separate task, like in approach (1) above. In additon, we create a global parameter set $\Theta$ that works as a prior to all the task specific parameters (or be regularized towards in a frequentist vocabulary).
The final posterior will look something like this:
%
\begin{equation} \label{eq:posterior}
    \sum_{n=1}^N  \sum_{i \in D_n} P(text_i | \theta_n) + \sum_{n=1}^N N(\theta_n | \Theta, \sigma)
\end{equation}

where
$D_n$ is the dataset corresponding to task $n$,
$P(text_i | \theta)$ is the likelihood of sentence $text_i$ evaluated on the parameters $\theta$,

@@

\section{Introduction}
\section{Related works}
\section{Method: Hierarchical LLM}
\section{Experiments}
\section{Conclusion}

>>>>

\section{Introduction}

\section{Related works}
\begin{itemize}
    \item explain what pretraining LLM means
    \item Explain how LORA fine tuned is done on each task
    \item Explain how LORA finetuned also could be done on all tasks together if they have a similar data distribution.
\end{itemize}

\section{Method: Hierarchical LLM}

\subsection{Task and Data Structure}
We define our study across $D$ distinct tasks, denoted as $d=1,2,...,D$. For each task, our dataset comprises $N_d$ documents, with each document containing a sequence of $W_n$ tokens. This dataset structure is pivotal for our analysis and can be formally represented as:

\begin{equation} \label{eq:data}
\mathcal{D} = \{{ w_{d,n,1:W_n} }{n=1}^{N_d} \}_{d=1}^D
\end{equation}

In this notation, $w_{d,n,1:W_n}$ represents the token sequence in the $n^{th}$ document of the $d^{th}$ task. For simplification in subsequent discussions, we will omit the indices $d$ and $n$ when their reference is evident from the context.

\subsection{Pretrained Autoregressive Language Model}
Our foundational model is a pretrained autoregressive language model, denoted by its parameters $\theta_{full}$ and its architecture, referred to as $LLM$. The model's probability distribution is given by:

\begin{equation} \label{eq:LLMprob}
P(w_{1:W} | \theta_{full}) = \prod_{i=1}^{W-1} LLM(w_{i+1} | w_{1:i})
\end{equation}
%
Here, $\theta_{full}$ represents the set of parameters obtained from pretraining on a large and diverse dataset, distinct from our current dataset $\mathcal{D}$.
We omit to write $\theta_{full}$ explicitly in the model, as it always depdends on it.

\subsection{The Hierarchical LORA finetuned}
For each task $d$ we have a LORA parameter set $\theta_d$ that is used to fine-tune to that task.
Hence, the likelihood of our problem follows from combining equation \ref{eq:data} and \ref{eq:LLMprob}:
\begin{equation} \label{eq:lik}
    \text{L}(\mathcal{D} | \theta_1, \theta_2, \ldots, \theta_D) = \prod_{d=1}^D \prod_{n=1}^{N_d} \prod_{i=1}^{W_n-1} \text{LLM}(w_{d,n,i+1} | w_{d,n,1:i}; \theta_d)
\end{equation}
%

The prior of

\begin{equation}
    P(\theta_{1:D}) = \prod_{d=1}^D N(\theta_d ; \Theta, \sigma I)
\end{equation}
%
where $\Theta$ is a set of hierarchical parameters, $I$ is a unit diagonal matrix and $\sigma \leq 0$ is a scalar hyperparameter setting how similar task parameters $\theta_d$ should be the hierarchical parameters $\Theta$.

\begin{itemize}
    \item Typically one finetune each task separately, as written in \ref{eq:lik}.
    \item If the D tasks are more similar to eachother than the pretraining corpus, the optimal parameters $\hat{\theta}_d$ should share some information during training in order to maximize their likelihood.
\end{itemize}

\section{Experimental setup}
We choose the `opt-350m` base model and construct three tasks that it has likely not seem as they are in Norwegian:
\begin{itemize}
    \item A subset of Norwegian parliament speeches of politicians containing 1000 examples
    \item A subset of a translation dataset from Norwegian bokmål to Norwegian Nynorsk (NbAiLab-nbnn-translation) containing 1000 examples
    \item A subset of a translation dataset from Norwegian bokmål to Norwegian Nynorsk (NbAiLab-nbnn-translation) containing 200 examples
\end{itemize}

These datasets will have a shared domain (Norwegian) that is different from what the pretrained foundational model is trained on.
We use two tasks containing the same dataset distribution to see how sharing knowledge through the hierarchical prior can benefit datasets with few examples.

\section{Conclusion}

@@

In addition to these two approaches, we present generalization of the two approaches: the Hierarchical Low-Rank Adaption.
Given a pretrained LLM and $n$ tasks, we construct $n$ sets of LoRA parameters $\theta_n$ that each are trained on its separate task, like in approach (1) above. In additon, we create a global parameter set $\Theta$ that works as a prior to all the task specific parameters (or be regularized towards in a frequentist vocabulary).
The final posterior will look something like this:
%
\begin{equation} \label{eq:posterior}
    \sum_{n=1}^N  \sum_{i \in D_n} P(text_i | \theta_n) + \sum_{n=1}^N N(\theta_n | \Theta, \sigma)
\end{equation}
%
where
$D_n$ is the dataset corresponding to task $n$,
$P(text_i | \theta)$ is the likelihood of sentence $text_i$ evaluated on the parameters $\theta$,

@@
